# Data analysis

## Load packages

```{r}
library(lattice)
library(tidyverse)
library(brms)
source("./r_utils/my_settings.R")
```

## Data

### Load data

```{r}
modals <- read.csv("./data/modals_data.csv", row.names=1)

# exclude USA
modals <- subset(modals, variety != "USA")
modals <- droplevels(modals)
str(modals)
```

### Inspect data

Distribution of observations across text files (extent of clustering)

```{r, fig.height=3, fig.width=3}
histogram(as.numeric(xtabs(~text_ID, data=modals)), nint=15)
```

Distribution of observations across conditions
```{r}
# for the tabular overview only: group source.ct (-1, -.5, 0, .5, 1) into 3 categories
modals$source_3 <- factor(ifelse(modals$source.ct < 0, "obj", ifelse(modals$source.ct == 0, "os", "subj")))

# Mode
ftable(modals[,c(4,1)])
```
```{r}
ftable(modals[,c(4,8,1)])
```

```{r}
ftable(modals[,c(4,8,11,1)])
```

### Visualization: Descriptive

Mode by variety
```{r fig.height=3, fig.width=4}
modals %>% group_by(variety, mode) %>% 
  dplyr::summarize(n_must = sum(verb=="must"),
                   n_have = sum(verb=="have to"),
                   n_need = sum(verb=="need to"),
                   n_got = sum(verb=="(have) got to")) %>% 
  mutate(n_total = n_must + n_have + n_need + n_got) %>% 
  mutate(p_must = n_must/n_total,
         p_have = n_have/n_total,
         p_need = n_need/n_total,
         p_got = n_got/n_total) %>% 
  select(variety, mode, p_must, p_have, p_need, p_got) %>% 
  gather(p_must, p_have, p_need, p_got, key="verb", value="share") %>% 
  ggplot(aes(x=mode, y=share, color=verb, group=verb)) + 
  geom_point() + geom_line() + facet_wrap(~ variety) +
  theme_classic()


modals %>% group_by(variety, mode, subject_3) %>% 
  dplyr::summarize(n_must = sum(verb=="must"),
                   n_have = sum(verb=="have to"),
                   n_need = sum(verb=="need to"),
                   n_got = sum(verb=="(have) got to")) %>% 
  mutate(n_total = n_must + n_have + n_need + n_got) %>% 
  mutate(p_must = n_must/n_total,
         p_have = n_have/n_total,
         p_need = n_need/n_total,
         p_got = n_got/n_total) %>% 
  select(variety, mode, subject_3, p_must, p_have, p_need, p_got) %>% 
  gather(p_must, p_have, p_need, p_got, key="verb", value="share") %>% 
  ggplot(aes(x=variety, y=share, color=verb, group=verb)) + 
  geom_point() + geom_line() + facet_wrap(~mode + subject_3) +
  theme_classic()
```

Grammatical subject by variety
```{r fig.height=3, fig.width=4}
modals %>% group_by(variety, subject_3) %>% 
  dplyr::summarize(n_must = sum(verb=="must"),
                   n_have = sum(verb=="have to"),
                   n_need = sum(verb=="need to"),
                   n_got = sum(verb=="(have) got to")) %>% 
  mutate(n_total = n_must + n_have + n_need + n_got) %>% 
  mutate(p_must = n_must/n_total,
         p_have = n_have/n_total,
         p_need = n_need/n_total,
         p_got = n_got/n_total) %>% 
  select(variety, subject_3, p_must, p_have, p_need, p_got) %>% 
  gather(p_must, p_have, p_need, p_got, key="verb", value="share") %>% 
  ggplot(aes(x=subject_3, y=share, color=verb, group=verb)) + 
  geom_point() + geom_line() + facet_wrap(~ variety) +
  theme_classic()
```

Source of authority by variety
```{r fig.height=3, fig.width=4}
modals %>% group_by(variety, source_3) %>% 
  dplyr::summarize(n_must = sum(verb=="must"),
                   n_have = sum(verb=="have to"),
                   n_need = sum(verb=="need to"),
                   n_got = sum(verb=="(have) got to")) %>% 
  mutate(n_total = n_must + n_have + n_need + n_got) %>% 
  mutate(p_must = n_must/n_total,
         p_have = n_have/n_total,
         p_need = n_need/n_total,
         p_got = n_got/n_total) %>% 
  select(variety, source_3, p_must, p_have, p_need, p_got) %>% 
  gather(p_must, p_have, p_need, p_got, key="verb", value="share") %>% 
  ggplot(aes(x=source_3, y=share, color=verb, group=verb)) + 
  geom_point() + geom_line() + facet_wrap(~ variety) +
  theme_classic()
```

# Fit regression models

## Construct weak priors based on literature

```{r}
p_have = .45
p_must = .30
p_got = .20
p_need = .05

log(p_have/p_must)
log(p_need/p_must)
log(p_got/p_must)

```


```{r}
str(modals)

# set reference level for outcome categories
modals$verb <- relevel(modals$verb, ref="must")

# create contrasts for grammatical subject (effect coding/deviation coding/sum contrasts)

#      subj_c1 subj_c2
# 1st       +1       0
# 2nd        0      +1
# 3rd       -1      -1

modals$subj_c1 <- ifelse(modals$subject_3 == "first", 1, ifelse(modals$subject_3 == "second", 0, -1))
modals$subj_c2 <- ifelse(modals$subject_3 == "first", 0, ifelse(modals$subject_3 == "second", 1, -1))

m_1 = 
  brm(verb ~ (subj_c1 + subj_c2 + source.ct + spoken.ct) * variety
      + (subj_c1 + subj_c2 + source.ct | text_ID),
      family = "categorical", data = modals,
      warmup=1000, iter= 2000, cores=3, chains=3,
      seed=1985,
      prior=c(
        set_prior("normal( -.4, 3)",  class = "Intercept", dpar="muhavegotto"),
        set_prior("normal(  .4, 3)",  class = "Intercept", dpar="muhaveto"),
        set_prior("normal(-1.8, 3)",  class = "Intercept", dpar="muneedto"),
        set_prior("normal(0, 2)", class = "b"),
        set_prior("exponential(1)", class = "sd", dpar = "muhaveto"),
        set_prior("exponential(1)", class = "sd", dpar = "muhavegotto"),
        set_prior("exponential(1)", class = "sd", dpar = "muneedto")
        )
      )

saveRDS(m_1, file = "C:/Users/ba4rh5/Local Folders/My Local Files/R models/consultation/schuetzler-herzky-modals/m_1.rds")
print(m_1)

saveRDS(c(0,0,0,0), "./models/test.rds")
saveRDS(c(0,0,0,0), "./models/test2.rds")
```

## Inspect fitted probabilities

```{r}
m_1 <- readRDS("C:/Users/ba4rh5/Local Folders/My Local Files/R models/consultation/schuetzler-herzky-modals/m_1.rds")

m_predicted <- predict(m_1)

d_check <- modals[,-c(5,6,7)]
d_check$p_hat_must <- m_predicted[,1]
d_check$p_hat_have <- m_predicted[,2]
d_check$p_hat_need <- m_predicted[,3]
d_check$p_hat_got <- m_predicted[,4]

plot1 <- histogram(~p_hat_must, data=d_check, par.settings=my_settings, axis=axis_L, breaks=seq(0,1,length.out=50), col="grey", ylim=c(0, 43))
plot2 <- histogram(~p_hat_have, data=d_check, par.settings=my_settings, axis=axis_L, breaks=seq(0,1,length.out=50), col="grey", ylim=c(0, 43))
plot3 <- histogram(~p_hat_need, data=d_check, par.settings=my_settings, axis=axis_L, breaks=seq(0,1,length.out=50), col="grey", ylim=c(0, 43))
plot4 <- histogram(~p_hat_got, data=d_check, par.settings=my_settings, axis=axis_L, breaks=seq(0,1,length.out=50), col="grey", ylim=c(0, 43))

print(plot1, position=c(0,0,.25,1), more=T)
print(plot2, position=c(.25,0,.5,1), more=T)
print(plot3, position=c(.5,0,.75,1), more=T)
print(plot4, position=c(.75,0,1,1))

```


# Model checking

## Residuals
```{r fig.height=4, fig.width=4}
d_check$resid_must <- d_check$p_hat_must - as.numeric(d_check$verb=="must") 
d_check$resid_have <- d_check$p_hat_have - as.numeric(d_check$verb=="have to") 
d_check$resid_need <- d_check$p_hat_need - as.numeric(d_check$verb=="need to") 
d_check$resid_got <- d_check$p_hat_got - as.numeric(d_check$verb=="(have) got to") 

xyplot(resid_must ~ source_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_have ~ source_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_need ~ source_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_got ~ source_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)

xyplot(resid_must ~ subject_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_have ~ subject_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_need ~ subject_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
xyplot(resid_got ~ subject_3 | mode + variety, data=d_check, jitter.x=TRUE, alpha=.3, col=1, ylim=c(-1,1), cex=.5)
```
## Compute posterior predictions
```{r}

```


